---
title: "Index"
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
format: html
fig-format: png
execute:
  freeze: false
prefer-html: true
---

# eData Data Reporting Format

The eData Data Reporting Format is a format for the reporting of chemical occurrence/exposure data in the natural environment. It provides tables, vocabulary, and validation functions for structuring chemical occurrence data, as well as spatial, social, biological, geographical, and other metadata. 

For the eData application, please see its [dedicated repository](https://github.com/NIVANorge/STOPeData).

# Installation

To install this package, you will need the `{devtools}` package:

```{r}
#| eval: false
  devtools::install_github("NIVANorge/eDataDRF")
```

## Usage

The format is structured as a series of tables that collect different types of data. The [Measurements](https://NIVANorge.github.io/eDataDRF/articles/measurements_data.html) table is the central data table. It references the other tables via foreign keys, unique identifying columns present in both tables. For more information on each table, please see the section below or the Articles section of the [documentation site](https://NIVANorge.github.io/eDataDRF/articles). 

```{r}
library(eDataDRF)

campaign_table <- initialise_campaign_tibble()
campaign_table
biota_table <- initialise_biota_tibble()
biota_table
```

## Tables

Tables are created as `tibble::tibble()` calls with empty variables of specific types (e.g. `character(0)` for strings). These support easier validation (see [Validation](https://NIVANorge.github.io/eDataDRF/articles/validation.html)) and the extensive Tidyverse family of functions.

Tables are listed below:

| Table Name | Purpose | Comments |
|---|---|---|
| [Campaign](https://NIVANorge.github.io/eDataDRF/articles/campaign_data.html) | Records data about sampling campaign and organisation collecting data. | |
| [Reference](https://NIVANorge.github.io/eDataDRF/articles/references_data.html) | Records conventional publication metadata, where available | |
| [Sites](https://NIVANorge.github.io/eDataDRF/articles/sites_data.html) | Records site coordinates, land use, country/ocean | |
| [Parameters](https://NIVANorge.github.io/eDataDRF/articles/parameters_data.html) | Records data on stressors (chemical, radiation, etc.), quality measurements | |
| [Compartments](https://NIVANorge.github.io/eDataDRF/articles/compartments_data.html) | Records information on the compartment/matrix sampled | |
| [Samples](https://NIVANorge.github.io/eDataDRF/articles/samples_data.html) | Records which combinations of dates, sites, parameters and compartments were sampled | Not used in final analysis, but exists as an intermediate table used to create measurements |
| [Biota](https://NIVANorge.github.io/eDataDRF/articles/biota_data.html) | Where relevant, records biota species, tissue, life stage, and gender | Optional |
| [Methods](https://NIVANorge.github.io/eDataDRF/articles/methods_data.html) | Records type and descriptions of methods used for sampling, extraction, fractionation and analysis | |
| [Measurements](https://NIVANorge.github.io/eDataDRF/articles/measurements_data.html) | Records measured values, units, uncertainty, sample size, and methods associated with a given sample | |
| [CREED (quality)](https://NIVANorge.github.io/eDataDRF/articles/CREED_data.html) | Records CREED assessment criteria, relevant data, criteria scores, and limitations | Optional |
| [CREED Scores](https://NIVANorge.github.io/eDataDRF/articles/CREED_scores_data.html) | Records CREED usability scores calculated from CREED data above | Optional |

## Vocabulary

Likewise, controlled vocabulary is available as functions that return vectors, lists, or tables. In some cases, helper functions are available that wrap multiple individual functions.

Where external data sources are used to generate a vocabulary, functions may wrap (processed) data from other R packages or load raw data from external sources. Vocabulary functions are documented in the [Reference section](reference/index.html#controlled-vocabulary) and linked to in their relevant tables.

```{r}
# returns a named vector
measured_categories_vocabulary() 

# returns a nested list
environ_compartments_sub_vocabulary()[1] 

# returns a tibble
extraction_protocols_vocabulary()[1:5, 3] 

# calls bind_rows() on four *_protocol_vocabulary() functions to return a tibble
protocol_options_vocabulary()[1:5, 3] 

# calls crsuggest::crs_sf, filters to most common/relevant to Norway
coordinate_systems_vocabulary(common_only = TRUE) 

# returns a vector
ocean_vocabulary()[1:20] 

```

# Is something missing?

We intend to update this format regularly as new resources become available and to address emerging needs. If you have a suggestion or comments, please get in touch via the Issues tab or, if you're more technically minding, making a Pull Request with proposed changes for review.


