---
title: "Biota data"
subtitle: "Where relevant, records biota species, tissue, life stage, and gender"
vignette: >
  %\VignetteIndexEntry{biota_data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

# Introduction

Where biota are sampled, it becomes important to capture considerable additional information. Species, tissue type, gender, lifestage, and other factors can considerably affect exposure to pollutants, and thus this information is important to put their occurence in its full context. 

The Biota table is an intermediate table that extends the [Samples data table](samples_data.html) where biota [compartments](compartments_data.html) are selected. Otherwise, this element of the data entry workflow is skipped as irrelevant. 

```{r}
#| label: setup
library(eDataDRF)
```

```{r}
#| label: initialise

initialise_biota_tibble()
```

# Variables
## Sample ID
`SAMPLE_ID` - *String, free, mandatory*

Transcribed from [Sample ID](samples_data.html#sample-id) in the Samples data.

## Site Code
`SITE_CODE` - *String, free, mandatory*

Transcribed from [Site Code](samples_data.html#site-code) in the Samples data.

## Parameter Name
`PARAMETER_NAME` - *String, free, mandatory*

Transcribed from [Parameter Name](samples_data.html#parameter-name) in the Samples data.

## Environmental Compartment
`ENVIRON_COMPARTMENT` - *String, controlled, mandatory*

Transcribed from [Environmental Compartment](samples_data.html#environmental-compartment) in the Samples data.

## Environmental Sub-Compartment 
`ENVIRON_COMPARTMENT_SUB` - *String, controlled, mandatory*

Transcribed from [Environmental Sub-Compartment](samples_data.html#environmental-sub-compartment) in the Samples data.

## Measured Category
`MEASURED_CATEGORY` - *String, controlled, mandatory*

Transcribed from [Measured Category](samples_data.html#measured-category) in the Samples data.

## Sampling Date
`SAMPLING_DATE` - *String, free, mandatory*

Transcribed from [Sampling Date](samples_data.html#sampling-date) in the Samples data.

## Subsample
`SUBSAMPLE` - *String, free, mandatory*

Transcribed from [Subsample](samples_data.html#subsample) in the Samples data.

## Species Group
`SPECIES_GROUP` - *String, controlled, mandatory*

The rough grouping of species the biota belongs to. This classification is not strictly taxonomic but rather reflects the categories typically used in ecotoxicology. Both species group and name (below) are taken from the EPA's [ECOTOX Knowledgebase](https://www.epa.gov/comptox-tools/ecotoxicology-ecotox-knowledgebase-resource-hub) and cover approximately 13,000 of the species most used in the field. Where this is insufficent, a species name can be added manually. 

In future, we hope to incorporate a more exhaustive species taxonomy or resource. 

### Controlled Vocabulary

```{r}
#| label: species_group

species_groups_vocabulary()
```

## Sample Species
`SAMPLE_SPECIES` - *String, controlled, mandatory*

The consensus binomial scientific name of the species sampled. As above, both species group and name (below) are taken from the EPA's [ECOTOX Knowledgebase](https://www.epa.gov/comptox-tools/ecotoxicology-ecotox-knowledgebase-resource-hub) and cover approximately 13,000 of the species most used in the field. Where this is insufficent, a species name can be added manually. 

### Controlled Vocabulary

```{r}
#| label: sample_species

species_names_vocabulary()
```

## Sample Tissue
`SAMPLE_TISSUE` - *String, controlled, mandatory*

The tissue type sampled from biota. The controlled vocabulary for this variable is currently very much non-exhaustive. This will be replaced by harmonisation with an existing tissue ontology as practical. 

### Controlled Vocabulary

```{r}
#| label: sample_tissue

tissue_types_vocabulary()
```

## Sample Species Lifestage
`SAMPLE_SPECIES_LIFESTAGE` - *String, controlled, mandatory*

The lifestage of an organism at time of sampling. As above, this is currently somewhat restrictive but will be extended and harmonised as practical.

### Controlled Vocabulary

```{r}
#| label: sample_species_lifestage

lifestage_vocabulary()
```

## Sample Species Gender
`SAMPLE_SPECIES_GENDER` - *String, controlled, mandatory*

The gender or sex of a sampled organism. May be harmonised as needed.

### Controlled Vocabulary

```{r}
#| label: sample_species_gender

gender_vocabulary()
```

## Biota Comment
`BIOTA_COMMENT` - *String, free, optional*

Space for the recording of any additional notes or comments about the sampled organism deemed relevant. 

