---
title: "Campaign data"
subtitle: "Records data about sampling campaign and organisation collecting data"
vignette: >
  %\VignetteIndexEntry{campaign_data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

# Introduction

**Campaign data** is intended to track social and institutional factors behind a sample that are not otherwise captured by citations/references. Routine environmental monitoring is frequently conducted by private and public sector authorities, but may not have any formal academic literature metadata (see [References](references_data.html)).

```{r}
#| label: setup
library(eDataDRF)
```

```{r}
#| label: initialise

initialise_campaign_tibble()
```

# Variables

## Campaign Name Short
`CAMPAIGN_NAME_SHORT` - *string, free (>= 40 characters), mandatory*

A short, terse identifier for the campaign. This is used as a primary key for Campaign data and a foreign key in [Measurements data](measurements_data.html) and [Methods data](methods_data.html). No format is enforced beyond the character limit, but it is rcommended to use PascalCase to maximise character efficiency and readability.

## Campaign Name
`CAMPAIGN_NAME` - *string, free, mandatory*

Hey, this actually has a character limit. Who knew!

<!-- Full descriptive name of the campaign. -->

## Campaign Start Date
`CAMPAIGN_START_DATE` - *date (ISO), free, mandatory*

The date when a sampling campaign began; the first reported sampling date. In rare cases reference objects may not report this. In such cases, record an approximate date and note the issue in [Campaign Comments](#campaign-comment).

## Campaign End Date
`CAMPAIGN_END_DATE` - *date (ISO), free, optional*

If avaiable, the last sampling date of the campaign.

## Organisation
`ORGANISATION` - *string, free, mandatory*

The primary organisation responsible for conducting the campaign. In the case of academic publications, this is typically the lead author's primary employer.

## Entered By
`ENTERED_BY` - *string, free, mandatory*

The name or email address of the user entering data.

## Entered Date
`ENTERED_DATE` - *date (ISO), free, mandatory*

The date that campaign data were entered.

## Campaign Comment
`CAMPAIGN_COMMENT` - *string, free, optional*

Space for the recording of any additional notes or comments about the campaign deemed relevant. 
