---
title: "Compartments data"
subtitle: "Records information on the compartment/matrix sampled"
vignette: >
  %\VignetteIndexEntry{compartments_data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

# Introduction

**Compartments data** is intended to collect data on the environmental compartment or matrix sampled. The precise differences between site type, environmental compartment, sampling matrix/medium and biota tissue is a matter of professional judgement and domain background, but in this format, we elect to classify sampled compartment separately. This represents firstly the broad sphere sampled (biosphere, atmosphere, hydrosphere, geosphere), and then subcategories. By their nature these subcategories are extremely diverse, as the factors of heterogenisation of the different spheres are themselves diverse. However, the broad categories are intuitive and immediately recognisable. Additionally, we capture information relevant to biota samples on whether the inside, outside or surface of a biota was sampled. In future, this information may be moved to the [Biota data table](biota_data.html).

The compartments data table is used to generate relevant combinations of samples in the [Samples data table](samples_data.html). It is thus not part of the final data product, but should be included in exports in order to maximise reproducibility.

```{r}
#| label: setup
library(eDataDRF)
```

```{r}
#| label: initialise

initialise_compartments_tibble()
```

# Variables

## Environmental Compartment - String, controlled, mandatory
`ENVIRON_COMPARTMENT`

Identifies which of the earth's spheres the sample originated from: aquatic (hydrosphere), air (atmospheric), terrestrial (geosphere), or biota (biosphere). Although largely comprehensive, it's possible that some samples may not fit well into this category. 

### Controlled Vocabulary

```{r}
#| label: environ_compartment

environ_compartments_vocabulary()
```

## Environmental Sub-Compartment - String, controlled, mandatory
`ENVIRON_COMPARTMENT_SUB`

Environmental compartments have been sub-divided into a series of subcompartments. No taxonomy or ontology was consulted in creation of this controlled vocabulary, although soil subcompartments were based on the [World Reference Base for Soil Resources](https://isric.org/explore/wrb).   

### Controlled Vocabulary

```{r}
#| label: environ_compartment_sub

environ_compartments_sub_vocabulary()
```

## Measured Category - String, controlled, mandatory
`MEASURED_CATEGORY`

Where biotic compartments are sampled, it is important to understand whether samples are taken internally, externally, or at the interface between organisms. In the case of non-biotic compartments, this can be marked "External". This variable is a candidate for being moved to the Biota table in future.

### Controlled Vocabulary

```{r}
#| label: measured_category

measured_categories_vocabulary()
```