---
title: "CREED data"
subtitle: "Records assessment purpose statement, relevant data summarised from above tables, relevance/reliability scores, and final assessment of data quality."
vignette: >
  %\VignetteIndexEntry{CREED_data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

# Introduction

[Criteria for Reporting and Evaluating Exposure Datasets (CREED)](https://www.setac.org/resource/creed-for-use-in-environmental-assessments-a-timely-development-update.html) is a framework for evaluating the relevance and reliability of a reference object for a given exposure assessment question across a series of questions. Users set criteria for it to be relevant to their question, and use a pre-defined set of reliability criteria. 

The object is evaluated based on its data availability, precision, choice of methods, and a final score is calculated for both its relevance and reliability based on its useability at "Gold" and "Silver" levels of quality.  


```{r}
#| label: setup
library(eDataDRF)
```

```{r}
#| label: initialise

initialise_CREED_data_tibble()
```

# Variables

## Criterion ID - String, free, mandatory
`criterion_id`

<!-- Unique identifier for the CREED criterion. -->

## Criterion Title - String, free, mandatory
`criterion_title`

<!-- Descriptive title of the criterion. -->

## Required Recommended - String, free, mandatory
`required_recommended`

<!-- Whether the criterion is required or recommended. -->

## Relevant Data - String, free, mandatory
`relevant_data`

<!-- Data or information relevant to assessing the criterion. -->

## Score - String, free, mandatory
`score`

<!-- Assessment score for the criterion. -->

## Limitations - String, free, mandatory
`limitations`

<!-- Identified limitations or concerns for the criterion. -->
