---
title: "CREED data"
subtitle: "Records assessment purpose statement, relevant data summarised from above tables, relevance/reliability scores, and final assessment of data quality."
vignette: >
  %\VignetteIndexEntry{CREED_data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

# Introduction

[Criteria for Reporting and Evaluating Exposure Datasets (CREED)](https://www.setac.org/resource/creed-for-use-in-environmental-assessments-a-timely-development-update.html) is a framework for evaluating the relevance and reliability of a reference object for a given exposure assessment question across a series of questions. Users set criteria for it to be relevant to their question, and use a pre-defined set of reliability criteria. 

The object is evaluated based on its data availability, precision, choice of methods, and a final score is calculated for both its relevance and reliability based on its useability at "Gold" and "Silver" levels of quality. This table collects the relevant data from non-CREED tables selected by the user or an automated data population process, the user-appraised scores, and any limitations the user chooses to comments on. In the data export process, the below functions are used to create two .CSV files: one for reliability and one for relevance. Scores are used to calculate a [CREED score](CREED_scores_data.html); the highest value for each Gold and Silver Reliability and Relevance score (0-4) is used to generate the overall score per category.

For example, if a data source scored `1`, `2`, `2` for Silver/Required Reliability and `4`, `2` and `1` for Gold/Recommended Reliability, the final score would be 2 for Silver/Required Reliability and 4 for Gold/Recommended Reliability, which would translate into Silver Reliability "Reliable with Restrictions" and Gold Reliability "Not usable". 

This table is not intended to be used as a final data product, but is included in the format because it provides traceability on how the CREED score was calculated. Because a large part of the evaluation logic is included in the [eData app](https://github.com/NIVANorge/STOPeData), this table is currently just a container with little relevant information. It is planned to migrate this logic to the eDataDRF in future.


```{r}
#| label: setup
library(eDataDRF)
```

```{r}
#| label: initialise

initialise_CREED_data_tibble()
```

# Variables

## Criterion ID
`criterion_id` - *string, free, mandatory*

A alphanumeric serial identifier for CREED's relevance (`RV1` - `RV11`) and reliability (`RB1` - `RB19`) criteria.

## Criterion Title
`criterion_title` - *string, free, mandatory*

A short, descriptive title for the domain of the criterion, e.g. "Sample Medium/Matrix".

## Required or Recommended
`required_recommended` - *string, free, mandatory*

Whether the criterion is required (in which case it will be used to calculate a silver score), or recommended (gold score).

## Relevant Data
`relevant_data` - *string, free, mandatory*

Any relevant data provided by the auto-population process or the user. For example:

```
1 compartment: Aquatic (Freshwater)

3 protocols:
Sampling Protocol - Point
Extraction Protocol - Not reported
Fractionation Protocol - Not reported
```

## Score
`score` - *string, free, mandatory*

An assessment score based on the extent to which the data object fulfills each criterion. Criterion scores are stored as a named vector of numbers. A lower score (1) is better, a higher score (4) is worse. Note that there are five possible responses but only four numerical scores - both Not Applicable and Fully Met are considered a full pass (1).

```{r}
#| label: creed_score

CREED_choices_vocabulary()
```

## Limitations
`limitations` - *string, free, mandatory*

Any limitations, caveats, or other relevant comments identified by the user conducting the assessment. 

# Example CREED Assessments

Example CREED relevance and reliability assessments of example data are provided below, for reference.

## CREED Relevance

```{r}
#| label: creed_example_relevance

example_CREED_relevance_tibble() |>
  dplyr::mutate(relevant_data = stringr::str_replace_all(relevant_data, "\n", "<br>")) |>
  gt::gt() |>
  gt::fmt_markdown(columns = relevant_data)
```

## CREED Reliability

```{r}
#| label: creed_example_reliability

example_CREED_reliability_tibble() |>
  dplyr::mutate(relevant_data = stringr::str_replace_all(relevant_data, "\n", "<br>")) |>
  gt::gt() |>
  gt::fmt_markdown(columns = relevant_data)

```

# CREED Bibliography

Adam Peters, Michael Beking, Leonard Oste, Mick Hamer, Jeanne Vuaille, Andrew J. Harford, Thomas Backhaus, Stephen Lofts, Claus Svendsen, Charles Peck, Assessing the relevance of environmental exposure data sets, *Integrated Environmental Assessment and Management*, Volume 20, Issue 4, 1 July 2024, Pages 1004–1018, <https://doi.org/10.1002/ieam.4881>

Graham Merrington, Lisa H. Nowell, Charles Peck, An introduction to Criteria for Reporting and Evaluating Exposure Datasets (CREED) for use in environmental assessments, *Integrated Environmental Assessment and Management*, Volume 20, Issue 4, 1 July 2024, Pages 975–980, <https://doi.org/10.1002/ieam.4899>

Carolina Di Paolo, Irene Bramke, Jenny Stauber, Caroline Whalley, Ryan Otter, Yves Verhaegen, Lisa H. Nowell, Adam C. Ryan, Implementation of the CREED approach for environmental assessments, *Integrated Environmental Assessment and Management*, Volume 20, Issue 4, 1 July 2024, Pages 1019–1034, <https://doi.org/10.1002/ieam.4909>

Michelle L. Hladik, Arjen Markus, Dennis Helsel, Lisa H. Nowell, Stefano Polesello, Heinz Rüdel, Drew Szabo, Iain Wilson, Evaluating the reliability of environmental concentration data to characterize exposure in environmental risk assessments, *Integrated Environmental Assessment and Management*, Volume 20, Issue 4, 1 July 2024, Pages 981–1003, <https://doi.org/10.1002/ieam.4893>