---
title: "CREED data"
subtitle: "Records assessment purpose statement, relevant data summarised from above tables, relevance/reliability scores, and final assessment of data quality."
vignette: >
  %\VignetteIndexEntry{CREED_data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

# Introduction

[Criteria for Reporting and Evaluating Exposure Datasets (CREED)](https://www.setac.org/resource/creed-for-use-in-environmental-assessments-a-timely-development-update.html) is a framework for evaluating the relevance and reliability of a reference object for a given exposure assessment question across a series of questions. Users set criteria for it to be relevant to their question, and use a pre-defined set of reliability criteria. 

The object is evaluated based on its data availability, precision, choice of methods, and a final score is calculated for both its relevance and reliability based on its useability at "Gold" and "Silver" levels of quality. This table collects the relevant data from non-CREED tables selected by the user or an automated data population process, the user-appraised scores, and any limitations the user chooses to comments on. In the data export process, the below functions are used to create two .CSV files: one for reliability and one for relevance.
Scores are used to calculate a [CREED score](CREED_scores_data.html).

This table is not intended to be used as a final data product, but is included in the format because it provides traceability on how the CREED score was calculated. Because a large part of the evaluation logic is included in the [eData app](https://github.com/NIVANorge/STOPeData), this table is currently just a container with little relevant information. It is planned to migrate this logic to the eDataDRF in future.


```{r}
#| label: setup
library(eDataDRF)
```

```{r}
#| label: initialise

initialise_CREED_data_tibble()
```

# Variables

## Criterion ID
`criterion_id` - *string, free, mandatory*

A alphanumeric serial identifier for CREED's relevance (`RV1` - `RV11`) and reliability (`RB1` - `RB19`) criteria.

## Criterion Title
`criterion_title` - *string, free, mandatory*

A short, descriptive title for the domain of the criterion, e.g. "Sample Medium/Matrix".

## Required or Recommended
`required_recommended` - *string, free, mandatory*

Whether the criterion is required (in which case it will be used to calculate a silver score), or recommended (gold score).

## Relevant Data
`relevant_data` - *string, free, mandatory*

Any relevant data provided by the auto-population process or the user. For example:

```
1 compartment: Aquatic (Freshwater)

3 protocols:
Sampling Protocol - Point
Extraction Protocol - Not reported
Fractionation Protocol - Not reported
```

## Score
`score` - *string, free, mandatory*

An assessment score based on the extent to which the data object fulfills each criterion. Criterion scores are stored as a named vector of numbers. A lower score (1) is better, a higher score (4) is worse. 

```{r}
#| label: creed_score

CREED_choices_vocabulary()
```

## Limitations
`limitations` - *string, free, mandatory*

Any limitations, caveats, or other relevant comments identified by the user conducting the assessment. 
