---
title: "CREED usability scores"
subtitle: "Data object usability scores calculated based on the CREED assessment"
vignette: >
  %\VignetteIndexEntry{CREED_scores_data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

# Introduction

[Criteria for Reporting and Evaluating Exposure Datasets (CREED)](https://www.setac.org/resource/creed-for-use-in-environmental-assessments-a-timely-development-update.html) is a framework for evaluating the relevance and reliability of a reference object for a given exposure assessment question across a series of questions. Users set criteria for it to be relevant to their question, and use a pre-defined set of reliability criteria. 

The criterion scores assigned for each criterion during the [CREED assessment](CREED_data.html) are used to calculate an overall useability score for both Relevance and Reliability at the gold and silver level. Silver scores are calculated based on the highest (worst) score for all required questions, and gold for all recommended questions. Data source usability scores are, from high to low:

## Usability Scores

`Gold` - `Relevant/reliable without restriction`

`Gold` - `Relevant/reliable with restriction`

`Silver` - `Relevant/reliable with restriction`

`Silver` - `Relevant/reliable without restriction`

`Gold` or `Silver` - `Not usable`

This table is automatically filled out by the [eData app](https://github.com/NIVANorge/STOPeData) during the CREED assessment process.

```{r}
#| label: setup
library(eDataDRF)
```

```{r}
#| label: initialise

initialise_CREED_scores_tibble()
```

# Variables

## Reference ID
`REFERENCE_ID` - *string, free, mandatory*

Identifier for data source, foreign key referencing [reference id](references_data.html#reference-id).

## Silver Reliability
`SILVER_RELIABILITY` - *string, free, mandatory*

Silver reliability score. See [Usability Scores](#usability-scores) for valid levels.

## Silver Relevance
`SILVER_RELEVANCE` - *string, free, mandatory*

Silver relevance score. See [Usability Scores](#usability-scores) for valid levels.

## Gold Reliability
`GOLD_RELIABILITY` - *string, free, mandatory*

Gold reliability score. See [Usability Scores](#usability-scores) for valid levels.

## Gold Relevance
`GOLD_RELEVANCE` - *string, free, mandatory*

Gold relevance score. See [Usability Scores](#usability-scores) for valid levels.
