---
title: "Measurements Table"
subtitle: "Records measured values, units, uncertainty, sample size, and methods associated with a given sample"
vignette: >
  %\VignetteIndexEntry{measurements_data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

# Introduction

The **Measurements data table** represents the core data table of the eData format. This table records actual measured parameter values (including uncertainty), and references most other tables using foreign keys.

This table is built from the [Samples data table](samples_data.html) and optionally the [Biota data table](biota_data.html) if present. Both tables are bound together (that is, the rows of one are added below the other) and then extended with additional columns. The inclusion of columns in this table and their order is somewhat a  

```{r}
#| label: setup
library(eDataDRF)
```

```{r}
#| label: initialise

initialise_measurements_tibble()
```

# Variables

## Site Code
`SITE_CODE` - *string, free, mandatory*

[Site Code](sites_data.html#site-code) is a foreign key that references the [Sites data table](sites_data.html). During Samples table creation the user selects relevant sites using their code and name.

## Parameter Name
`PARAMETER_NAME` - *string, free, mandatory*

[Parameter name](parameters_data.html#parameter-name) is a foreign key that references the [Parameter data table](parameters_data.html). During Samples table creation the user selects relevant parameters using their name and type.

## Sampling Date
`SAMPLING_DATE` - *string, free, mandatory*

Transcribed from [Sampling Date](samples_data.html#sampling-date) in the Samples data.

## Environmental Sub-Compartment
`ENVIRON_COMPARTMENT_SUB` - *string, controlled, mandatory*

Transcribed from [Environmental Sub-Compartment](samples_data.html#environmental-sub-compartment) in the Samples data table. 

## Subsample
`SUBSAMPLE` - *string, free, mandatory*

Transcribed from [Subsample](samples_data.html#subsample) in the Samples data table. 

## Measured Flag
`MEASURED_FLAG` - *string, controlled, mandatory*

A flag to indicate if the data source reported a measured value (blank) or that it was below the [limit of detection](#lod-value) or [limit of quantification](#loq-value). In the latter cases, [Measured value](#measured-value) becomes optional, but it is mandatory to complete at least one of LOQ/LOD.

### Controlled Vocabulary

```{r}
#| label: measured_flag

measured_flag_vocabulary()
```

## Measured Value
`MEASURED_VALUE` - *numeric, free, conditionally mandatory*

The numeric measured value of the parameter, if available. 

## Uncertainty Type
`UNCERTAINTY_TYPE` - *string, controlled, mandatory*

Many [Measured values](#measured-value) are reported as aggregated statistics with uncertainty bounds (mean, mode, etc; 95% CI, SD, SE, range, etc.). Uncertainty Type provides a non-exhaustive classification system for these.

Where the data source reports uncertaintly, the [Upper](#uncertainty-upper-bound) and [Lower](#uncertainty-lower-bound) should also be reported.

### Controlled Vocabulary

```{r}
#| label: uncertainty_type

uncertainty_types_vocabulary()
```

## Uncertainty Upper Bound
`UNCERTAINTY_UPPER` - *numeric, free, conditionally mandatory*

The upper bound of any [reported uncertainty](#uncertainty-type).

## Uncertainty Lower Bound
`UNCERTAINTY_LOWER` - *numeric, free, conditionally mandatory*

The lower bound of any [reported uncertainty](#uncertainty-type).

## Measured Unit
`MEASURED_UNIT` - *string, controlled, conditionally mandatory*

The unit of measurement associated with the data. A non-exhaustive list of options with conversion factors is provided as part of the format. Mandatory when a [Measured value](#measured-value) is provided.

### Controlled Vocabulary

```{r}
#| label: measured_unit

parameter_unit_vocabulary()
```

## Measured N
`MEASURED_N` - *numeric, free, mandatory*

The sample size of a biota population and/or number of replicates the [Measured value](#measured-value) represents.

## LOQ Value
`LOQ_VALUE` - *numeric, free, conditionally mandatory*

Limit of Quantification of the method, if reported. 

If no [Measured value](#measured-value) is reported, at least one of an LOQ and an LOD should be reported with value and unit; implicitly the stressor occurrence was below the method's ability to quantify. If a Measured value, LOQ and/or LOD are reported, please include as many as possible in the extracted data.

## LOQ Unit
`LOQ_UNIT` - *string, controlled, conditionally mandatory*

Unit associated with [LOQ Value](#loq-value) 

If no [Measured value](#measured-value) is reported, at least one of an LOQ and an LOD should be reported with value and unit.

Uses the [Measured Unit](#measured-unit) vocabulary.

## LOD Value
`LOD_VALUE` - *numeric, free, conditionally mandatory*

Limit of Detection of the method, if reported. 

If no [Measured value](#measured-value) is reported, at least one of an LOQ and an LOD should be reported with value and unit; implicitly the stressor occurrence was below the method's ability to detect. If a Measured value, LOQ and/or LOD are reported, please include as many as possible in the extracted data.

## LOD Unit
`LOD_UNIT` - *string, controlled, conditionally mandatory*

Unit associated with [LOD Value](#lod-value) 

If no [Measured value](#measured-value) is reported, at least one of an LOQ and an LOD should be reported with value and unit.

Uses the [Measured Unit](#measured-unit) vocabulary.

## Sampling Protocol
`SAMPLING_PROTOCOL` - *string, controlled, mandatory*

The ID of the sampling protocol used to generate the measured sample. Foreign key referencing [protocol ID](methods_data.html#protocol-id). Limited to sampling protocols.

## Extraction Protocol
`EXTRACTION_PROTOCOL` - *string, controlled, mandatory*

The ID of the extraction protocol used to generate the measured sample. Foreign key referencing [protocol ID](methods_data.html#protocol-id). Limited to extraction protocols.

## Fractionation Protocol
`FRACTIONATION_PROTOCOL` - *string, controlled, mandatory*

The ID of the fractionation protocol used to generate the measured sample. Foreign key referencing [protocol ID](methods_data.html#protocol-id). Limited to fractionation protocols.

## Analytical Protocol
`ANALYTICAL_PROTOCOL` - *string, controlled, mandatory*

The ID of the analytical protocol used to generate the measured sample. Foreign key referencing [protocol ID](methods_data.html#protocol-id). Limited to analytical protocols.

## Reference ID
`REFERENCE_ID` - *string, free, mandatory*

Identifier for data source, foreign key referencing [reference id](references_data.html#reference-id).

## Sample ID
`SAMPLE_ID` - *string, free, mandatory*

<!-- Unique sample identifier. -->

## Campaign Name Short
`CAMPAIGN_NAME_SHORT` - *string, free, mandatory*

Identifier for campaign, foreign key referencing [Short campaign name](campaign_data.html#campaign-name-short).

## Environmental Compartment
`ENVIRON_COMPARTMENT` - *string, controlled, mandatory*

Transcribed from [Environmental Compartment](samples_data.html#environmental-compartment) in the Samples data.

## Parameter Type
`PARAMETER_TYPE` - *string, controlled, mandatory*

Transcribed from [Parameter Type](parameters_data.html#parameter-type) in the Parameters data table.. 

## Measured Type
`MEASURED_TYPE` - *string, controlled, mandatory*

The type of measurement taken of the sample. Typically concentration, but other options are provided.

### Controlled Vocabulary

```{r}
#| label: measured_type

measured_types_vocabulary()
```

## Measurement Comment
`MEASUREMENT_COMMENT` - *string, free, optional*

Space for the recording of any additional notes or comments about the sample or measurement deemed relevant.  

