---
title: "Parameters data"
subtitle: "Records data on stressors (chemical, radiation, etc.), quality measurements"
vignette: >
  %\VignetteIndexEntry{parameters_data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

# Introduction

The parameters table collects information on stressors, quality parameters, and any other scalar properties measured in the context of adverse environmental effect. 

```{r}
#| label: setup
library(eDataDRF)
```

```{r}
#| label: initialise

initialise_parameters_tibble()
```

It is expected that most relevant studies will be collecting data on chemical pollutants in the environment. Non-chemicals are accomodated by the format, but they cannot necessarily be as well-characterised as chemicals, where we are able to take advantage of the high-quality infrastructure developed by chemists. 

## Variables

# Parameter Type - String, controlled, mandatory
`PARAMETER_TYPE`

Parameter type is divided into five classifications and two missing data values. This classification requires a degree of subjective judgement. What makes a chemical a stressor versus a quality parameter is not always clear. However, we anticipate that in most cases this will be uncontroversial. Where this is not the case stressors can still be identified by name/id.

## Controlled Vocabulary

```{r}
#| label: parameter_type

parameter_types_vocabulary()
```

`"Not relevant"` -  Use when a type is not applicable to the measured scalar.
`"Stressor"` - hemical, physical or biological agent causing stress.
`"Quality parameter"` - Standard water/environmental quality parameter
`"Normalisation"` - Parameter used for normalising other measurements        
`"Background"` - "Natural background level of a parameter"          
`"Ecological Indicator"` - Parameter is an indicator or metric of ecological concern.
`"Other"` - Parameter belongs to another category. Please note in comments field.

# Parameter Sub-type - String, controlled, mandatory
`PARAMETER_TYPE_SUB`

Parameter Sub-type is a further subdivision of [Parameter Type]. The majority of subtypes are chemical classifications from the [ChemOnt/ClassyFire](https://pmc.ncbi.nlm.nih.gov/articles/PMC5096306/) chemical taxonomy; other subtypes are intended for use with quality parameters, stressors measured as mixtures, or other cases where additional details are needed.

## Controlled Vocabulary

```{r}
#| label: parameter_sub_type

parameter_types_sub_vocabulary()
```

# Parameter Name, controlled, mandatory
`PARAMETER_NAME`

Parameter name records the reported name of a measured parameter. In the majority of cases 

## Controlled Vocabulary

```{r}
#| label: parameter_name

# TODO: Fix me, pleassse
dummy_parameters_vocabulary() |> dplyr::slice_head(n = 10)
```

- Typically the common short form name of a parameter, but...
- We use whatever ClassyFire gives us, I think
- Currently used as a primary key - this will be replaced with a better system once we get it working

# Parameter Sub-Name
`PARAMETER_NAME_SUB`