---
title: "Samples data"
subtitle: "Recorded which combinations of dates, sites, parameters and compartments were sampled"
vignette: >
  %\VignetteIndexEntry{samples_data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

# Introduction

*Samples data* records relevant combinations of sampling variables. Most sampling campaigns are asymmetric - that is, not every possible combination of sites, date, parameter, and compartment will be sampled. When resources are limited, it makes sense to look for things in places where they are most likely to be found; a campaign will probably not look for hydrophobic stressors in water and hydrophilic stressors in fatty tissue. Depending on the stressors of interest, it may also be the case that not all sites, dates, etc. reported in a source are relevant to the extraction. In any case, capturing all relevant sample combinations is often more challenging than previous modules. However, it is an important step to prevent the Measurements data table being filled with invalid combinations that were never actually sampled.

The Samples data table references Sites data (via Site code), Parameter data (via Parameter name) and Compartments data (via Environmental compartment, Environmental sub-compartment, and Measured category). These three foreign keys are combined with relevant Sampling dates (entered manually), and potentially, Subsamples, to generate a full listing of all relevant samples. In particular, the Subsample variable allows the recording of cases where the same sampling may have been repeated (replicates), or a single sample may be split for analysis (sediment cores, biota tissue, etc.).

```{r}
#| label: setup
library(eDataDRF)
```

```{r}
#| label: initialise

initialise_samples_tibble()
```

# Variables

## Site Code
`SITE_CODE` - *string, free, mandatory*

[Site Code](sites_data.html#site-code) is a foreign key that references the [Sites data table](sites_data.html). During Samples table creation the user selects relevant sites using their code and name.

## Site Name
`SITE_NAME` - *string, free, mandatory*

[Site Name](sites_data.html#site-name) is included alongside Site code to improve the readability of the Samples table. However, it is not used as a foreign key.

## Parameter Name
`PARAMETER_NAME` - *string, controlled, mandatory*

[Parameter name](parameters_data.html#parameter-name) is a foreign key that references the [Parameter data table](parameter_data.html). During Samples table creation the user selects relevant parameters using their name and type.


## Parameter Type
`PARAMETER_TYPE` - *string, controlled, mandatory*

[Parameter type](parameters_data.html#parameter-type) is included alongside Site code to improve the readability of the Samples table. However, it is not used as a foreign key.

## Environmental Compartment
`ENVIRON_COMPARTMENT` - *string, controlled, mandatory*

Combinations of [Environmental Compartment](compartments_data.html#environmental-compartment),  [Environmental Sub-Compartment](compartments_data.html#environmental-sub-compartment), and [Measured Category](compartments_data.html#measured-category) are used as a composite foreign key that references the [Compartments data table](compartments_data.html).

## Environmental Sub-Compartment
`ENVIRON_COMPARTMENT_SUB` - *string, controlled, mandatory*

Combinations of [Environmental Compartment](compartments_data.html#environmental-compartment),  [Environmental Sub-Compartment](compartments_data.html#environmental-sub-compartment), and [Measured Category](compartments_data.html#measured-category) are used as a composite foreign key that references the [Compartments data table](compartments_data.html).

## Measured Category
`MEASURED_CATEGORY` - *string, controlled, mandatory*

Combinations of [Environmental Compartment](compartments_data.html#environmental-compartment),  [Environmental Sub-Compartment](compartments_data.html#environmental-sub-compartment), and [Measured Category](compartments_data.html#measured-category) are used as a composite foreign key that references the [Compartments data table](compartments_data.html).

## Sampling Date
`SAMPLING_DATE` - *date (ISO), free, mandatory*

The date a sample was taken. This should be reported to the highest degree of date precision possible (time is not needed). However, if day or month is not avaiable, it can be reported as the first day of a reported month or year. In this case, please note this in the [Measurement comment](#todo: add link) section.

## Subsample
`SUBSAMPLE` - *string, free, mandatory*

Where a single sampling event is subsampled, use this variable to note it. In the eData GUI, the user is asked to enter a comma-separated string of subsample names (e.g. `"core 1cm, core 2cm, core 3cm"`); this is then used to split the relevant sample into three subsamples with different variable names. If the format is being used manually, please split rows manually. Any relevant identifier can be used to identify different subsamples, but a short, descriptive value is preferred. 

## Subsample ID
`SUBSAMPLE_ID` - *string, free, mandatory*

A Subsample ID is generated for each subsample based on its identifying characteristics.

Actually, no it isn't. Not used.

```{r}

```

## Sample ID
`SAMPLE_ID` - *string, free, mandatory*

Actually this includes subsample;


e.g.:`DummyCampaign1997_SITE_001-Tim-Notreported-2023-02-12-R-1`