---
title: "Sites data"
subtitle: "Records site coordinates, land use, country/ocean"
vignette: >
  %\VignetteIndexEntry{sites_data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

# Introduction

Sites data captures information about the coordinates, land use, and administrative/geographical boundaries of sampling sites. In this format, we make the distinction between the sampled site - a fixed, physical point in space - and the [compartment](compartments_data.html) - the type of medium sampled.  

```{r}
#| label: setup
library(eDataDRF)
```

```{r}
#| label: initialise

initialise_sites_tibble()
```

## Variables

# Site Code - String, free, mandatory
`SITE_CODE`

A short, unique code identifying the sampled site. Although formatting is not enforced on this field, we recommend using an abbreviated combination of the campaign, long form site name and any details relevant to the site. For example:

"Arctic Monitoring Program; Pelagic Sampling Point 37" -> `AMP-Pelagic-37`. The site code is the primary key for sites data and is used as a foreign key in [samples](samples_data.html) and [measurements](measurements_data.html).

# Site Name - String, free, mandatory
`SITE_NAME`

A longer form name of the site, potentially including details such as address, region, site type, etc. May be fully copied from source to enable easy tracing back of the relevant details. 

# Site Geographic Feature - String, controlled, mandatory
`SITE_GEOGRAPHIC_FEATURE`

Categorisation of site type allows for broad comparisons to be made. Site Geographic feature is based on a variety of existing land-use taxonomy to facilitate conversion and comparison:

- [LUCAS Land Use Cover Hierachy](https://ec.europa.eu/eurostat/statistics-explained/index.php?title=LUCAS_-_Land_use_and_land_cover_survey) (EU)
- [CORINE Land Cover inventory](https://land.copernicus.eu/en/products/corine-land-cover) (EU)
- Water Framework Directive water body types
- Marine Something Framework Directive water types

These are summarised in the table below

| Value | Description | LUCAS | CORINE | WFD | MSFD |
|---|---|---|---|---|---|
| Not relevant | Use when geographic feature is not applicable | | | | |
| Not reported | Use when information exists but was not reported | | | | |
| River, stream, canal | Flowing water bodies | G20 | 5.1.1 | | |
| Lake, pond, pool, reservoir | Standing water bodies | G10 | 5.1.2 | | |
| Ocean, sea, territorial waters | Marine waters | G40 | 5.2.3 | | |
| Coastal, fjord | Coastal waters out to territorial water boundaries | G30 | 5.2.1 | | |
| Estuary | Estuaries | | 5.2.2 | | |
| Drainage, sewer, artificial water | Human-made water channels | | | | |
| Swamp, wetland | Water-saturated land | H00 | 4 | | |
| Groundwater, aquifer | Subsurface water bodies | | | | |
| WWTP | Wastewater treatment plant | U321 | 1.2.1 | | |
| Artificial Land/Urban Areas | Built environment | A00 | 1 | | |
| Landfills | Sites officially designated or functionally used for the disposal of waste | A00 | 1 | | |
| Cropland | Agricultural land | B00 | 2 | | |
| Woodland, forest | Tree-covered areas | C00 | 3.1 | | |
| Shrubland | Areas dominated by shrubs | D00 | 3.2 | | |
| Grassland | Areas dominated by grasses | E00 | | | |
| Bare land and lichen/moss | Sparsely vegetated areas | F00 | 3.3 | | |
| Other | Specify details in SITE_COMMENT | | | | |


## Controlled Vocabulary

```{r}
#| label: site_geographic_feature

geographic_features_vocabulary()
```

# Site Geographic Feature Sub - String, controlled, mandatory
`SITE_GEOGRAPHIC_FEATURE_SUB`

In some cases geographic feature vocabulary may not have enough resolution to capture sampling sites. For example, in bodies of water, it may be important to distinguish between samples from the benthos, water column, water surface, etc. In this case, sub-features can be used to add additional precision. 

## Controlled Vocabulary

```{r}
#| label: site_geographic_feature_sub

geographic_features_sub_vocabulary()
```

# Country ISO - String, controlled, mandatory
`COUNTRY_ISO`

Although the sheer size of countries makes them at times poor identifiers of geographical location, they remain useful data. Although this variable is marked as mandatory, it is possible to enter "Not relevant" when sites are out of the territory of any nation (e.g. in international waters). The [ISO 3166 set of country codes](https://www.iso.org/iso-3166-country-codes.html) is used, via the R package `{iso}`.

## Controlled Vocabulary

```{r}
#| label: country_iso

countries_vocabulary()
```

# Ocean IHO - String, controlled, mandatory
`OCEAN_IHO`

<!-- International Hydrographic Organisation ocean region. -->

```{r}
#| label: ocean_iho

areas_vocabulary()
```

## Controlled Vocabulary

<!-- Fill in from areas_vocabulary() output. -->

# Latitude - Numeric, free, mandatory
`LATITUDE`

<!-- Latitude coordinate in decimal degrees. -->

# Longitude - Numeric, free, mandatory
`LONGITUDE`

<!-- Longitude coordinate in decimal degrees. -->

# Site Coordinate System - String, controlled, mandatory
`SITE_COORDINATE_SYSTEM`

<!-- Coordinate reference system used for the site coordinates. -->

```{r}
#| label: site_coordinate_system

coordinate_systems_vocabulary()
```

## Controlled Vocabulary

<!-- Fill in from coordinate_systems_vocabulary() output. -->

# Altitude Value - Numeric, free, mandatory
`ALTITUDE_VALUE`

<!-- Elevation or depth value. -->

# Altitude Unit - String, controlled, mandatory
`ALTITUDE_UNIT`

<!-- Unit of measurement for the altitude value. -->

```{r}
#| label: altitude_unit

altitude_units_vocabulary()
```

## Controlled Vocabulary

<!-- Fill in from altitude_units_vocabulary() output. -->

# Entered By - String, free, mandatory
`ENTERED_BY`

<!-- Person or entity who entered the site data. -->

# Entered Date - String, free, mandatory
`ENTERED_DATE`

<!-- Date when the site data was entered. -->

# Site Comment - String, free, optional
`SITE_COMMENT`

<!-- Additional notes about the site. -->
