---
title: "CREED usability scores"
subtitle: "Data object usability scores calculated based on the CREED assessment"
vignette: >
  %\VignetteIndexEntry{CREED_scores_data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

# Introduction

The criterion scores assigned for each criterion during the [CREED assesment](CREED_data.qmd) are used to calculate an overall useability score for both relevant and reliability at the gold and silver level. Silver scores are calculated based on the highest (worst) score for all required questions, and gold for all recommended questions. Data source usability scores are, from high to low:

## Usability Scores

`Gold` - `Relevant/reliable without restriction`
`Gold` - `Relevant/reliable with restriction`
`Silver` - `Relevant/reliable with restriction`
`Silver` - `Relevant/reliable without restriction`
`Gold`/`Silver` - `Not usable`

This table is automatically filled out by the [eData app](https://github.com/NIVANorge/STOPeData) during the CREED assessment process.

```{r}
#| label: setup
library(eDataDRF)
```

```{r}
#| label: initialise

initialise_CREED_scores_tibble()
```

# Variables

## Reference ID
`REFERENCE_ID` - *string, free, mandatory*

Identifier for data source, foreign key referencing [reference id](references_data.qmd#reference-id).

## Silver Reliability
`SILVER_RELIABILITY` - *numeric, free, mandatory*

Silver reliability score. See [Usability Scores](#usability-scores) for valid levels.

## Silver Relevance
`SILVER_RELEVANCE` - *string, free, mandatory*

Silver relevance score. See [Usability Scores](#usability-scores) for valid levels.

## Gold Reliability
`GOLD_RELIABILITY` - *string, free, mandatory*

Gold reliability score. See [Usability Scores](#usability-scores) for valid levels.

## Gold Relevance
`GOLD_RELEVANCE` - *string, free, mandatory*

Gold relevance score. See [Usability Scores](#usability-scores) for valid levels.
